Analysing Constraint Grammars with a SAT-solver

### 2015-10-25 paper/lrec/eval_futurework.tex
nitpicking

\begin{table}[]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
                      & \textsc{nld}  & \textsc{spa}  & \textsc{fin}  \\ \hline
\# rules      & 59              & 279      & 1185     \\ \hline
\# readings      & 276             & 2191     & 2012    \\ \hline
\# conflicts       & 1               & 1        & \todo{n}    \\ \hline
\clock{} all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
\clock{} last rule       & 4s              & 7s                & 3min 40s    \\ \hline
% This is with just 2000ish variables for Finnish, below is with 5700

@@

\paragraph{Correctness} 
The test revealed problems in all grammars, and for the smaller
grammars, we were able to verify manually that the detected rules were
true negatives.
%The results for the Finnish grammar are inconclusive, 
% For the Finnish grammar, we did not have time to
% investigate all of them, but among those we could, we found both true and false negatives.
We did not check for false positives in any of the grammars.

@@

For the Finnish grammar, we could verify a number of tagset errors. 
The grammar was originally written in 1995, and converted to the
Omorfi tagset by \cite{pirinen2015}. For example, the Omorfi tagset
marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
grammar uses the old convention \texttt{PL2} in the rules.
As a result, all rules with possessive suffixes were detected as
impossible.

Because of performance issues, we ignored the fully specified readings from the
lexicon and used only tag combinations from the grammar:
for example, the rule \texttt{REMOVE (verb sg) IF (-1 det)}
gives us ``verb sg'' and ``det''.
The shortcut works most of the time, but it is possible to
miss some cases: e.g. \texttt{SELECT (verb sg) IF (0 p3)} 
assumes ``verb sg p3'', but our method only gives
``verb sg'' and ``p3'' separately. 
Some of the detected rules seem to be false negatives caused by this shortcut. 
Another suspect is our naive implementation of rules with \verb!*! in
their context.

@@

\paragraph{Performance} The speed of the check is an obvious issue,
especially for the Finnish grammar. 
As mentioned before, we used 2000 readings extracted from the grammar.
With the full set of 6000 readings from the lexicon, the speed
was even worse: checking the last rule took 30 minutes, as opposed to
3 minutes with 2000 readings. 
%We never finished running the whole rule set with 6000 readings, but
%it would have taken potentially days.
We plan to make an effort in optimising the code; currently it is just
a naive proof-of-concept implementation.

@@

%We will also improve usability, by finding out which rules contrad
For longer-term goals, we want to handle the full expressivity of CG-3,
with \textsc{map}, \textsc{add} and \textsc{substitute} rules, and
dependency structure.
As soon as the tools are mature enough, we want to
evaluate them with actual grammar writers,
in comparison with a corpus-based method or machine learning.
Finally, if the method proves feasible for CG, we want
to try applying it to other grammar formalisms.

>>>>

\begin{table}[]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
                      & \textsc{nld}  & \textsc{spa}  & \textsc{fin}  \\ \hline
\# rules      & 59              & 279      & 1185     \\ \hline
\# readings      & 276             & 2191     & 2012    \\ \hline
\# conflicts       & 1               & 1        & 51    \\ \hline
\clock{} all rules       & 7s              & 3min 22s    & 3h 26min    \\ \hline
\clock{} last rule       & 4s              & 7s                & 3min 40s    \\ \hline
% This is with just 2000ish variables for Finnish, below is with 5700

@@

\paragraph{Accuracy} 
The test revealed problems in all grammars, and for the smaller
grammars, we were able to verify manually that the detected rules were
true positives.
%The results for the Finnish grammar are inconclusive, 
% For the Finnish grammar, we did not have time to
% investigate all of them, but among those we could, we found both true and false negatives.
We did not check for false negatives in any of the grammars.

@@

For the Finnish grammar, we did not have time to investigate all the
results. We could verify approximately \nicefrac{1}{3} as true positive and
\nicefrac{1}{3} as false positive.
The verified true positives were all rule-internal conflicts. 
The following rule requires the target to be genitive and unambiguously nominative: 
\begin{itemize}
\item[]\texttt{"oma" SELECT GEN IF [...] (0C NOM) ;}
\end{itemize}
We also detected a number of tagset errors, . 
The grammar was originally written in 1995, and converted to the
Omorfi tagset by \cite{pirinen2015}, but there were 15 rules that used the old convention for possessive suffixes.

%  For example, the Omorfi tagset
% marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
% grammar uses the old convention \texttt{PL2} in the rules.
% As a result, 15 rules with possessive suffixes were detected as
% impossible.
To improve performance, we ignored the fully specified readings from the
lexicon and used only tag combinations from the grammar:
for example, the rule \texttt{REMOVE (verb sg) IF (-1 det)}
gives us ``verb sg'' and ``det''.
The shortcut works most of the time, but it is possible to
miss some cases: e.g. \texttt{SELECT PRON + REL IF (0 NOM)} 
assumes ``pron rel nom'', but our  only gives
``pron rel'' and ``nom'' separately. 
17 false positives were caused by this shortcut. 
In addition, 2 false positives were caused by the naive implementation of
rules with \verb!*! in their context.

@@

\paragraph{Performance} The speed of the check is an obvious issue,
especially for the Finnish grammar. 
As mentioned before, we used 2000 readings extracted from the grammar.
With the full set of 6000 readings from the lexicon, the speed
was even worse: checking the last rule took 30 minutes, as opposed to
3 minutes with 2000 readings. 
%We never finished running the whole rule set with 6000 readings, but
%it would have taken potentially days.
We plan to optimise the code; currently it is just
a naive proof-of-concept implementation.

@@

As for longer-term goals, we want to handle the full expressivity of CG-3,
with \textsc{map}, \textsc{add} and \textsc{substitute} rules, and
dependency structure. This also means finding different kinds of conflicts.
As soon as the tools are mature enough, we want to
evaluate them with actual grammar writers,
in comparison with a corpus-based method or machine learning.
Finally, if the method proves feasible for CG, we want
to try applying it to other grammar formalisms.
